h51495
s 00014/00015/00031
d D 1.3 10/05/12 19:24:01 starck 3 2
c update olivier
e
s 00029/00025/00017
d D 1.2 10/02/18 18:31:30 starck 2 1
c Update doc
e
s 00042/00000/00000
d D 1.1 10/02/18 09:30:29 starck 1 0
c date and time created 10/02/18 09:30:29 by starck
e
u
U
f e 0
t
T
I 1

D 2
\newpage
E 2
I 2
\newpage
E 2
\thispagestyle{empty}
$ $
\newpage

D 3
{\Huge \bf Acknowledgments}\label{forewd}\\
D 2
\vspace{1cm}
E 2
I 2
E 3
I 3
{\Huge \bf Acknowledgments}
\label{forewd}\\
E 3
\vspace{1cm}
E 2
\markboth{Acknowledgments}{Acknowledgments}
D 2
\addcontentsline{toc}{chapter}{Acknowledgments}

The authors of this software package would like to express their 
gratitude to all of their colleagues who have made this work  
possible through various contributions. 
We are especially thankful to our direct collaborators with whom we 
have been working for years on problems in statistics and data 
analysis in astrophysics and cosmology: Nabila Aghanim, Jacques Delabrouille, David Donoho, 
Olivier Forni, Jiashun Jin, Mai Nguyen and Guillaume Patanchon.
This package is a compilation of the algorithms and methods which were 
developed and/or used successfully in a number of applications 
reported in the following publications: \\
E 2
I 2
\addcontentsline{toc}{chapter}{Acknowledgments}

Main developers for this package are  O.~Fourt, J.-L. Starck, J.~Bobin, Y.~Moudden and  P.~Abrial.
The authors of this software package would like to express their 
gratitude to all of their colleagues who have made this work  
possible through various contributions. 
We are especially thankful to our direct collaborators with whom we 
have been working for years on problems in statistics and data 
analysis in astrophysics and cosmology: Nabila Aghanim, Jacques Delabrouille, David Donoho, Jalal Fadili, 
Olivier Forni, Jiashun Jin, Mai Nguyen and Guillaume Patanchon.
This package is a compilation of the algorithms and methods which were 
developed and/or used successfully in a number of applications 
reported in the following publications: \\
E 2
\begin{itemize}
D 2
\item[$\bullet$]{ \textbf{ Multi--Detector Multi--Component spectral matching and applications for CMB data analysis}, J. Delabrouille, J.-F. Cardoso and G. Patanchon, \emph{Monthly Notices of the Royal Astronomical Society}, 2003, vol. 346, n. 4, pp. 1089-1102 }
\item[$\bullet$]{ \textbf{ Detecting Cosmological non-Gaussian Signatures by Multi-scale Methods}, J.-L. Starck and N. Aghanim and O. Forni, \emph{Monthly Notices of the Royal Astronomical Society}, 2004, vol. 416, n. 4, pp. 9-17 }
E 2
I 2
D 3
\item[$\bullet$] { \textbf{ Blind Source Separation: The Sparsity Revolution},  J. Bobin, J.-L. Starck, M.J. Fadili, and Y. Moudden,  \emph{Advances in Imaging and Electron Physics }, Vol 152, pp 221 -- 306, 2008.  }
\item[$\bullet$] {  \textbf{ CMB Data Analysis and Sparsity},  P. Abrial, Y. Moudden, J.L. Starck,M.J. Fadili, J. Delabrouille, and M. Nguyen,   \emph{Statistical Methodology}, Vol 5, No 4, pp 289-298, 2008.}
\item[$\bullet$]{   \textbf{SZ and CMB reconstruction using GMCA}, J. Bobin, Y. Moudden,J.-L. Starck, M.J. Fadili, and N. Aghanim,  \emph{Statistical Methodology}, Vol 5, No 4, pp 307-317, 2008. }
\item[$\bullet$]{   \textbf{ Sparsity, Morphological Diversity and Blind Source Separation},  J. Bobin, J.-L. Starck, M.J. Fadili, and Y. Moudden, \emph{IEEE Transactions on Image Processing}, Vol 16, No 11, pp 2662-2674, 2007. }
\item[$\bullet$] {  \textbf{Morphological Component Analysis and Inpainting on the Sphere: Application in Physics and Astrophysics},   P. Abrial, Y. Moudden, J.L. Starck, J. Bobin, M.J. Fadili, B. Afeyan and M. Nguyen,  \emph{Journal of Fourier Analysis and Applications  (JFAA)}, 13, 6, pp 729-748, 2007. }
\item[$\bullet$]{   \textbf{  Morphological Component Analysis: an adaptative thresholding strategy},  J. Bobin, J.-L. Starck J. Fadili, Y. Moudden and D.L Donoho,  \emph{IEEE Transactions on Image Processing}, Vol 16, No 11, pp 2675--2681, 2007. }
\item[$\bullet$]{\textbf{ Detecting Cosmological non-Gaussian Signatures by Multi-scale Methods}, J.-L. Starck and N. Aghanim and O. Forni, \emph{Monthly Notices of the Royal Astronomical Society}, 2004, Vol. 416, n. 4, pp. 9-17 }
E 2
\item[$\bullet$]{ \textbf{ Cosmological Non-Gaussian Signatures Detection: Comparison of Statistical Tests}, J. Jin,  J.-L. Starck,  D.L. Donoho,  N. Aghanim and O. Forni, \emph{EURASIP Journal of Applied Signal Processing},  vol. 2005, n.15, pp 2470-2485, 2005.}
\item[$\bullet$]{ \textbf{ Blind Component Separation in Wavelet Space: Application to {CMB} Analysis}, Y. Moudden and J.-F. Cardoso and J.-L. Starck and J. Delabrouille, 
\emph{EURASIP Journal of Applied Signal Processing}, vol. 2005, n. 15, pp 2437-2454, 2005.}
D 2
\item[$\bullet$]{ \textbf{ Wavelets, Ridgelets and Curvelets on the Sphere},  J.-L. Starck,  P. Abrial, Y. Moudden and M. Nguyen, \emph{Astronomy and Astrophysics}, 2005, to appear}
\item[$\bullet$]{ \textbf{ Sunyaev-Zeldovich cluster reconstruction in multiband  bolometer camera surveys}, S. Pires, J.-B. Juin, D. Yvon, Y. Moudden, S. Anthoine and E. Pierpaoli, submitted to \emph{Astronomy and Astrophysics}, June 2005, available at: \\
 http://arxiv.org/pdf/astro-ph/0508641}\\
E 2
I 2
\item[$\bullet$]{ \textbf{ Wavelets, Ridgelets and Curvelets on the Sphere},  J.-L. Starck,  P. Abrial, Y. Moudden and M. Nguyen, \emph{Astronomy and Astrophysics}, vol. 446, pp 1191-1204, 2, 2006.}
\item[$\bullet$]{ \textbf{ Sunyaev-Zeldovich cluster reconstruction in multiband  bolometer camera surveys}, S. Pires, J.-B. Juin, D. Yvon, Y. Moudden, S. Anthoine and E. Pierpaoli, submitted to \emph{Astronomy and Astrophysics}, vol. 455,  pp.741-755, 2006.}
E 2
E 3
I 3
\item[$\bullet$] {\textbf{Blind Source Separation: The Sparsity Revolution}, J. Bobin, J.-L. Starck, M.J. Fadili, and Y. Moudden, \textit{Advances in Imaging and Electron Physics }, Vol 152, pp 221 -- 306, 2008.}
\item[$\bullet$] {\textbf{CMB Data Analysis and Sparsity}, P. Abrial, Y. Moudden, J.L. Starck,M.J. Fadili, J. Delabrouille, and M. Nguyen, \textit{Statistical Methodology}, Vol 5, No 4, pp 289-298, 2008.}
\item[$\bullet$] {\textbf{SZ and CMB reconstruction using GMCA}, J. Bobin, Y. Moudden,J.-L. Starck, M.J. Fadili, and N. Aghanim, \textit{Statistical Methodology}, Vol 5, No 4, pp 307-317, 2008.}
\item[$\bullet$] {\textbf{Sparsity, Morphological Diversity and Blind Source Separation}, J. Bobin, J.-L. Starck, M.J. Fadili, and Y. Moudden, \textit{IEEE Transactions on Image Processing}, Vol 16, No 11, pp 2662-2674, 2007.}
\item[$\bullet$] {\textbf{Morphological Component Analysis and Inpainting on the Sphere: Application in Physics and Astrophysics}, P. Abrial, Y. Moudden, J.L. Starck, J. Bobin, M.J. Fadili, B. Afeyan and M. Nguyen,  \textit{Journal of Fourier Analysis and Applications (JFAA)}, 13, 6, pp 729-748, 2007.}
\item[$\bullet$] {\textbf{Morphological Component Analysis: an adaptative thresholding strategy}, J. Bobin, J.-L. Starck J. Fadili, Y. Moudden and D.L Donoho, \textit{IEEE Transactions on Image Processing}, Vol 16, No 11, pp 2675--2681, 2007.}
\item[$\bullet$] {\textbf{Detecting Cosmological non-Gaussian Signatures by Multi-scale Methods}, J.-L. Starck and N. Aghanim and O. Forni, \textit{Monthly Notices of the Royal Astronomical Society}, 2004, Vol. 416, n. 4, pp. 9-17}
\item[$\bullet$] {\textbf{Cosmological Non-Gaussian Signatures Detection: Comparison of Statistical Tests}, J. Jin, J.-L. Starck, D.L. Donoho, N. Aghanim and O. Forni, \textit{EURASIP Journal of Applied Signal Processing}, vol. 2005, n.15, pp 2470-2485, 2005.}
\item[$\bullet$] {\textbf{Blind Component Separation in Wavelet Space: Application to {CMB} Analysis}, Y. Moudden and J.-F. Cardoso and J.-L. Starck and J. Delabrouille, \textit{EURASIP Journal of Applied Signal Processing}, vol. 2005, n. 15, pp 2437-2454, 2005.}
\item[$\bullet$] {\textbf{Wavelets, Ridgelets and Curvelets on the Sphere}, J.-L. Starck, P. Abrial, Y. Moudden and M. Nguyen, \textit{Astronomy and Astrophysics}, vol. 446, pp 1191-1204, 2, 2006.}
\item[$\bullet$] {\textbf{Sunyaev-Zeldovich cluster reconstruction in multiband bolometer camera surveys}, S. Pires, J.-B. Juin, D. Yvon, Y. Moudden, S. Anthoine and E. Pierpaoli, submitted to \textit{Astronomy and Astrophysics}, vol. 455,  pp.741-755, 2006.}
E 3
\end{itemize}
D 2

Other colleagues we would like to acknowledge include:
Bedros Afeyan, Jerome Bobin, Emmanuel Candes, Pierre-Fran{\c{c}}ois Honor\'e, Ludovic Poupard, 
Philippe Querre, Sandrine Pires, Ryad Sehil and Patricio Vielva.
E 2
I 2
Other colleagues we would like to acknowledge include:
D 3
Bedros Afeyan,  Emmanuel Candes, Pierre-Fran{\c{c}}ois Honor\'e, Ludovic Poupard, 
Philippe Querre, Sandrine Pires, Ryad Sehil and Patricio Vielva.
E 2
E 3
I 3
Bedros Afeyan, Emmanuel Candes, Pierre-Fran{\c{c}}ois Honor\'e, Ludovic Poupard, Philippe Querre, Sandrine Pires, Ryad Sehil and Patricio Vielva.
E 3

D 2
\newpage
E 2
I 2
\newpage
E 2
\thispagestyle{empty}
$ $
\newpage

D 2

E 1
E 2
I 2

E 2
