
\newpage
\thispagestyle{empty}
$ $
\newpage

{\Huge \bf Acknowledgments}\label{forewd}\\
\vspace{1cm}
\markboth{Acknowledgments}{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

The authors of this software package would like to express their 
gratitude to all of their colleagues who have made this work  
possible through various contributions. 
We are especially thankful to our direct collaborators with whom we 
have been working for years on problems in statistics and data 
analysis in astrophysics and cosmology: Nabila Aghanim, Jacques Delabrouille, David Donoho, 
Olivier Forni, Jiashun Jin, Mai Nguyen and Guillaume Patanchon.
This package is a compilation of the algorithms and methods which were 
developed and/or used successfully in a number of applications 
reported in the following publications: \\
\begin{itemize}
\item[$\bullet$]{ \textbf{ Multi--Detector Multi--Component spectral matching and applications for CMB data analysis}, J. Delabrouille, J.-F. Cardoso and G. Patanchon, \emph{Monthly Notices of the Royal Astronomical Society}, 2003, vol. 346, n. 4, pp. 1089-1102 }
\item[$\bullet$]{ \textbf{ Detecting Cosmological non-Gaussian Signatures by Multi-scale Methods}, J.-L. Starck and N. Aghanim and O. Forni, \emph{Monthly Notices of the Royal Astronomical Society}, 2004, vol. 416, n. 4, pp. 9-17 }
\item[$\bullet$]{ \textbf{ Cosmological Non-Gaussian Signatures Detection: Comparison of Statistical Tests}, J. Jin,  J.-L. Starck,  D.L. Donoho,  N. Aghanim and O. Forni, \emph{EURASIP Journal of Applied Signal Processing},  vol. 2005, n.15, pp 2470-2485, 2005.}
\item[$\bullet$]{ \textbf{ Blind Component Separation in Wavelet Space: Application to {CMB} Analysis}, Y. Moudden and J.-F. Cardoso and J.-L. Starck and J. Delabrouille, 
\emph{EURASIP Journal of Applied Signal Processing}, vol. 2005, n. 15, pp 2437-2454, 2005.}
\item[$\bullet$]{ \textbf{ Wavelets, Ridgelets and Curvelets on the Sphere},  J.-L. Starck,  P. Abrial, Y. Moudden and M. Nguyen, \emph{Astronomy and Astrophysics}, 2005, to appear}
\item[$\bullet$]{ \textbf{ Sunyaev-Zeldovich cluster reconstruction in multiband  bolometer camera surveys}, S. Pires, J.-B. Juin, D. Yvon, Y. Moudden, S. Anthoine and E. Pierpaoli, submitted to \emph{Astronomy and Astrophysics}, June 2005, available at: \\
 http://arxiv.org/pdf/astro-ph/0508641}\\
\end{itemize}

Other colleagues we would like to acknowledge include:
Bedros Afeyan, Jerome Bobin, Emmanuel Candes, Pierre-Fran{\c{c}}ois Honor\'e, Ludovic Poupard, 
Philippe Querre, Sandrine Pires, Ryad Sehil and Patricio Vielva.

\newpage
\thispagestyle{empty}
$ $
\newpage


