\documentclass[11pt,a4paper]{article}
\usepackage{psfig}
\usepackage{amstex}
 
\newcommand{\mw}{/export/home/starck/Main/sadam/tex/report/doc_mr2}
% \newcommand{\psdir}{/home/starck/Main/sun/starck/tex/report/FIG}
\newcommand{\psdir}{/home/starck/Main/sun/starck/tex/report/FIG}

\psfigurepath{\psdir/mr1:\psdir/demo:\psdir/cours:\psdir/edge:\psdir/compress:\psdir/fractal:\psdir/detect:\psdir/book98:\psdir/entropy:\psdir/iso:\psdir/cours}

\def\aa{{A\&A } }
\def\aj{{AJ} }
\def\apj{{ApJ\ }}
\def\va{{ Vistas in Astronomy}}
\def\pasp{{ PASP\ }}

\def\projmra{MR}
\def\defprojmra{The Multiresolution Analysis Software}

\medskip

\def\proj{MR/1 }
\def\defproj{The Multiresolution Environment}

\def\projmw{MR/2 }
\def\defprojmw{Multiscale Entropy and Applications}

\newcommand{\be}{\begin{eqnarray}}
\newcommand{\ee}{\end{eqnarray}}
\newcommand{\bn}{\begin{enumerate}}
\newcommand{\en}{\end{enumerate}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

% \makeindex
% \input psfig.tex

% \voffset -1.5cm
% \hoffset -1.cm
\evensidemargin 1.truecm
\oddsidemargin  1.truecm

 
\begin{document}
 18 transforms are implemented, which are grouped into 5 classes 
\begin{itemize}
\item{Class 1:} no decimation (transforms 1 to 6 and 10 to 13).
\item{Class 2:} pyramidal transform (transforms 7 to 9).
\item{Class 3:} orthogonal transform (14 and 15).
\item{Class 4:} Wavelet packets (16 and 17).
\item{Class 5:} Wavelet packets via the \`a trous algorithm  (18).
\end{itemize}

Morlet, Mexican hat,
and French hat wavelet transforms are non-dyadic (the resolution is
not decreased by a factor two between two scales), 
and 12 voices (fixed value) are calculated (instead of one for the
dyadic case) when the resolution is divided by two. Then the number of 
rows in the output image will be equal to 12*{\em number\_of\_scales}.
The Morlet wavelet is complex, so the wavelet transform is complex too.
Four transforms are non-redundant: the bi-orthogonal,
the lifting scheme, and the wavelet packet methods (transforms 16 and 17). 
In this case, the output is not an 
image but a signal (i.e.\ 1D rather than 2D), 
which has the same size as the original one.
 
\begin{enumerate}
\item linear wavelet transform: \`a trous algorithm 
\item B$_1$-spline wavelet transform: \`a trous algorithm 
\item B$_3$-spline wavelet transform: \`a trous algorithm 
\item derivative of a B$_3$-spline: \`a trous algorithm 
\item undecimated Haar wavelet transform: \`a trous algorithm
\item morphological median transform 
\item pyramidal linear wavelet transform 
\item pyramidal B$_3$-spline wavelet transform 
\item pyramidal median transform 
\item Morlet's wavelet transform \\
Continuous wavelet transform with a complex wavelet 
which can be decomposed into two parts, one for
the real part, and the other for the imaginary part:
\begin{eqnarray*}
\psi_r(x) & = & \frac{1}{\sqrt{2 \pi}\sigma}  e^{-\frac{x^2}{2\sigma^2}} \cos(2\pi\nu_0 {x \over \sigma}) \nonumber \\
\psi_i(x) & = & \frac{1}{\sqrt{2 \pi}\sigma}  e^{-\frac{x^2}{2\sigma^2}} \sin(2\pi\nu_0 {x \over \sigma}) 
\end{eqnarray*}
First scale $\sigma_0$ is chosen equal to $2 \nu_0$ and $\nu_0 = 0.8$, 
using 12 voices per octave.
\item Mexican hat wavelet transform \\
Continuous wavelet transform with the Mexican hat wavelet.
This is the second derivative of a Gaussian
\begin{eqnarray}
\psi(x) = (1 - {x^2 \over \sigma^2}) e^{-\frac{ x^2}{2\sigma^2}} 
\end{eqnarray}
First scale $\sigma_0$ is chosen equal to ${1 \over \sqrt{3}}$, using 12 voices per octave.
\item French hat wavelet transform \\
Continuous wavelet transform with the French  hat wavelet
\begin{eqnarray}
\psi(x) =  \left \{
\begin{array}{ll}
     0  & \mbox{ if }  \mid {x \over \sigma} \mid > 1 \\
     -1 & \mbox{ if } \mid  {x \over \sigma} \mid \in [{1 \over 3},1]  \\
     2  & \mbox{ if } \mid  {x \over \sigma} \mid < {1 \over 3}
\end{array}
\right.
\end{eqnarray}  
First scale $\sigma_0$ equal to $0.66$, and 12 voices per octave. 
\item Gaussian derivative wavelet transform \\
 Continuous wavelet transform. The wavelet is the first 
derivative of a Gaussian
 \begin{eqnarray}
 \psi(x) =  - x e^{-\frac{1}{2}x^2} 
 \end{eqnarray}
 First scale $\sigma_0$ equal to ${1 \over \sqrt{3}}$, and 12 voices per octave. 
\item (bi-) orthogonal transform. \\
Implemented filters are
\begin{enumerate}
\item Antonini 7/9 filters 
\item Daubechies filter 4 
\item Haar filter 
\item Odegard 7/9 filters
\end{enumerate}
\item (bi-) orthogonal transform via lifting scheme. \\
 Implemented lifting scheme method are:
\begin{enumerate}
\item  CDF WT 
\item  median prediction
\item  integer Haar WT 
\item  integer CDF WT 
\item  integer (4,2) interpolating transform 
\end{enumerate}}
\item Wavelet packets.
\item Wavelet packets via lifting scheme. 
\item Wavelet packets using the \`a trous algorithm.
\end{enumerate}
\end{itemize}


% \bibliographystyle{plain}
% \bibliography{starck,wave,restore,compress,iso,reg,ima,edge,fractal}

\end{document}
