
\newpage
\thispagestyle{empty}
$ $
\newpage

{\Huge \bf Acknowledgments}
\label{forewd}\\
\vspace{1cm}
\markboth{Acknowledgments}{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}


\section*{{\isap} }
{\isap} is a collection of packages, in IDL and C++.  The {\isap} environment has been designed by F. Sureau, F. Lanusse and 
J.-L. Starck. The authors of the different packages are:


\section*{\projsparse2d package}
Main developers for this package are  J.-L. Starck, A. Woiselle, J. Fadili, J. Bobin, and F. Sureau. 

This package is based on algorithms and methods described in: 
\begin{itemize}
\item[$\bullet$] {\textbf{Sparse Image \& Signal Processing: wavelets, curvelets, morphological diversity}, J.-L. Starck , F. Murtagh and J. Fadili, \textit{Cambridge University Press},  2010.}
\end{itemize}

\section*{\projmsvst package}
Main developers for this package are  B. Zhang, J. Fadili and J.-L. Starck. 

This package is based on methods described in: 
\begin{itemize}
\item[$\bullet$]  {\textbf{Wavelets, Ridgelets and Curvelets for Poisson Noise Removal}, B. Zhang, M.J. Fadili and J.-L. Starck,  
\textit{IEEE Transactions on Image Processing}, Vol 17, No 7, pp 1093--1108, 2008. }.
\end{itemize}

\section*{\projmrs package}
Main developers for this package are  F. Sureau, J.-L. Starck, J.~Bobin.   
Developers of previous versions are O.~Fourt, P.~Abrial and Y.~Moudden.
The authors of this software package would like to express their 
gratitude to all of their colleagues who have made this work  
possible through various contributions. 
We are especially thankful to our direct collaborators with whom we 
have been working for years on problems in statistics and data 
analysis in astrophysics and cosmology: Nabila Aghanim, Fabio Noviello,  Jacques Delabrouille, David Donoho, Jalal Fadili, 
Olivier Forni, Jiashun Jin and Mai Nguyen.
The original paper describing wavelets, ridgelets and curvelets on the sphere is:
\begin{itemize}
\item[$\bullet$] {\textbf{Wavelets, Ridgelets and Curvelets on the Sphere}, J.-L. Starck, P. Abrial, Y. Moudden and M. Nguyen, \textit{Astronomy and Astrophysics}, vol. 446, pp 1191-1204, 2, 2006.}
\end{itemize}
A more detailed description can be found in:
\begin{itemize}
\item[$\bullet$] {\textbf{Sparse Image \& Signal Processing: wavelets, curvelets, morphological diversity}, J.-L. Starck , F. Murtagh and J. Fadili, \textit{Cambridge University Press},  2010.}
\end{itemize}
Other colleagues we would like to acknowledge include:
Bedros Afeyan, Emmanuel Candes, Jean-Fran{\c{c}}ois  Cardoso,  Pierre-Fran{\c{c}}ois Honor\'e,  Guillaume Patanchon, Ludovic Poupard, Philippe Querre, Sandrine Pires, Ryad Sehil and Patricio Vielva.

This work was partially  supported by the French National Agency for Research (ANR -08-EMER-009-01) and by the European Research Council  (ERC-228261).

\section*{\projpol package}
Main developers for this package are  F. Sureau, J.-L. Starck, J.~Bobin. 
Developers of previous versions are O.~Fourt, P.~Abrial and Y.~Moudden.
The authors of this software package would like to express their 
gratitude to David Donoho, Morteza Shahram, and Inam Ur-Raman 

This package is based on algorithms and methods which were published in:
\begin{itemize}
\item[$\bullet$] {\textbf{Polarized Wavelets and Curvelets on the Sphere}, J.-L. Starck, Y. Moudden and J. Bobin, \textit{Astronomy and Astrophysics}, Vol 497, pp 931--943, 2009.}
\end{itemize}

This work was partially supported by the European Research Council  (ERC-228261).

\section*{\projmsvsts package}

Main developers for this package are  J.~Schmitt , J.-L. Starck and J. Fadili.
 
This package is based on algorithms and methods which were 
developed and/or used successfully in a number of applications 
reported in the following publications: \\
\begin{itemize}
\item[$\bullet$] {\textbf{Poisson Denoising on the Sphere: Application to the Fermi Gamma Ray Space Telescope},   J. Schmitt, J.L. Starck, J.M. Casandjian, J. Fadili, I. Grenier, \textit{Astronomy and Astrophysics},  517, A26, 2010.}
\item[$\bullet$] {\textbf{Multichannel Poisson Denoising and Deconvolution on the Sphere : Application to the Fermi Gamma Ray Space Telescope}, J. Schmitt, J.L. Starck, J.M. Casandjian, J. Fadili, I. Grenier,, \textit{Astronomy and Astrophysics}, 546, id.A114, pp10, 2012.}
\end{itemize}

This work was partially supported by the European Research Council  (ERC-228261).

\section*{\projsgal package}
Main developers for this package are  F.X.  Dup\'e, A. Rassat, A. Leonard, D. Machado and  J.-L. Starck.
This package is based on algorithms and methods which were published in: \\
\begin{itemize}
\item[$\bullet$] { \textbf{Darth Fader: Using wavelets to obtain accurate redshifts of spectra at very low signal-to-noise}, 
D. Machado, A. Leonard, J-L. Starck, and F. Abdalla,  \textit{Astronomy and Astrophysics}, 560, id.A83, pp 20, 2013.}
\item[$\bullet$] {\textbf{Measuring the Integrated Sachs-Wolfe Effect", Astronomy and Astrophysics}, F. -X Dupe, A. Rassat, A., J.-L. Starck, M.J. Fadili, \textit{Astronomy and Astrophysics}, 534, A51+, 2011.}
\end{itemize}
This work was partially supported by the European Research Council  (ERC-228261).

\section*{\projscmb package}
Main developers for this package are  P. Paykari, F. Lanusse, F. Sureeau, J. Bobin,  A. Rassat, J. Fadili and  J.-L. Starck.
This package is based on algorithms and methods which were published in: \\
\begin{itemize}
\item[$\bullet$] \textbf{On Preferred Axes in WMAP Cosmic Microwave Background Data after Subtraction of the Integrated Sachs-Wolfe Effect}, A. Rassat and J-L. Starck,  \textit{Astronomy and Astrophysics} , 557, id.L1, pp 7, 2013.
\item[$\bullet$] \textbf{Removal of two large scale Cosmic Microwave Background anomalies after subtraction of the Integrated Sachs Wolfe effect}, A. Rassat, J-L. Starck, and F.X. Dupe,  \textit{Astronomy and Astrophysics} , 557, id.A32, pp 15, 2013.
\item[$\bullet$] \textbf{WMAP 9-year CMB estimation using sparsity}, \textit{Astronomy and Astrophysics} , J. Bobin, F. Sureau, P. Paykari, A. Rassat, S. Basak and J.-L. Starck, 553, L4, pp 10, 2013.
\item[$\bullet$] \textbf{Sparsity and the Bayesian Perspective}, J.-L. Starck, D.L. Donoho, M.J. Fadili and A. Rassat,  \textit{Astronomy and Astrophysics} , 552, A133, 2013.
\item[$\bullet$]\textbf{Sparse component separation for accurate CMB map estimation} ,  J. Bobin, J.-L. Starck, F. Sureau and S. Basak,  \textit{Astronomy and Astrophysics} , 550, A73, 2013.
\item[$\bullet$]  \textbf{Low-l CMB Analysis and Inpainting}, \textit{Astronomy and Astrophysics} , J.-L. Starck, A. Rassat, and M.J. Fadili, 550, A15, 2013.
\item[$\bullet$]  \textbf{True CMB Power Spectrum Estimation}, P. Paykari, J.L. Starck, and J. Fadili,  \textit{Astronomy and Astrophysics} , 541, A74, 2012.
\item[$\bullet$] \textbf{CMB map restoration}, J. Bobin, J.-L. Starck, F. Sureau, and J. Fadili, , \textit{Advances in Astronomy} , 2012, Id703217, 2012.
\end{itemize}
This work was partially supported by the European Research Council  (ERC-228261).


\newpage
\thispagestyle{empty}
$ $
\newpage


